# (PART\*) Regression Models {-} 

# Measuring Performance in Regression Models

```{r chapter-05-startup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/")
library(knitr)
library(tidymodels)
library(patchwork)
```

tidymodels focuses more on data frames rather than individual vectors. Let's keep our observed and predicted values from a regression model in a data frame: 

```{r chapter-05-example-data}
library(tidymodels)
tidymodels_prefer()

numeric_predictions <- 
 tibble::tribble(
  ~observed, ~predicted,
       0.22,       0.24,
       0.83,       0.78,
      -0.12,      -0.66,
       0.89,       0.53,
      -0.23,        0.7,
       -1.3,      -0.75,
      -0.15,      -0.41,
       -1.4,      -0.43,
       0.62,       0.49,
       0.99,       0.79,
      -0.18,      -1.19,
       0.32,       0.06,
       0.34,       0.75,
       -0.3,      -0.07,
       0.04,       0.43,
      -0.87,      -0.42,
       0.55,      -0.25,
       -1.3,      -0.64,
      -1.15,      -1.26,
        0.2,      -0.07
 )
```

Let's visualize these values in a scatter plot, add a diagonal reference line, and scale the axes to be the same. We'll also add a residual plot and then put them in the same figure using the `r pkg(patchwork)` package:

```{r chapter-05-fig-03}
library(patchwork)
theme_set(theme_bw())

obs_vs_pred <- 
 ggplot(numeric_predictions, aes(x = observed, y = predicted)) + 
 geom_point() + 
 # Adds a diagonal reference line
 geom_abline(col = "darkgrey", lty = 2) + 
 # Scales the axes to be the same based on the range of both
 # dimensions. 
 coord_obs_pred()

residual_plot <- 
 numeric_predictions %>% 
 mutate(residual = observed - predicted) %>% 
 ggplot(aes(x = predicted, y = residual)) + 
 geom_point() + 
 geom_hline(yintercept = 0, col = "darkgrey", lty = 2) 

obs_vs_pred + residual_plot + plot_layout(ncol = 2)
```

The `r pkg(yardstick)` package contains functions for computing performance metrics. For example, for $R^2$ and RMSE: 

```{r chapter-5-basic}
numeric_predictions %>% 
 rsq(truth = observed, estimate = predicted)

numeric_predictions %>% 
 rmse(truth = observed, estimate = predicted)
```

These can also be combined into a _metric set_. A new function that produces multiple metrics is created: 

```{r chapter-5-metric-set}
reg_stats <- metric_set(rsq, rmse)

numeric_predictions %>% 
 reg_stats(truth = observed, estimate = predicted)
```


For a list of performance metrics in `r pkg(yardstick)`, see the [reference page](https://yardstick.tidymodels.org/reference/index.html). 

