# Case Study: Compressive Strength of Concrete Mixtures

```{r chapter-10-startup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/")
library(tidymodels)
library(workflowsets)
library(baguette)
library(rules)
library(janitor)
library(finetune)

cores <- parallel::detectCores()
if (!grepl("mingw32", R.Version()$platform)) {
 library(doMC)
 registerDoMC(cores = cores)
} else {
  library(doParallel)
  cl <- makePSOCKcluster(cores)
  registerDoParallel(cl)
}
```

This case study examines a large number of models that predict a property of concrete mixtures. The analysis here will be similar to the _APM_ approach but will illustrate two techniques: using workflow sets to launch models and efficient triage of models using race methods. This chapter uses a lot of R packages: `r pkg_text(c("tidymodels", "rules", "Cubist", "workflowsets", "AppliedPredictiveModeling", "glmnet", "nnet", "earth", "kernlab", "kknn", "baguette", "rpart", "ranger", "xgboost"))`.

There are two versions of the data. Like _APM_, we'll use the version where the concrete ingredients are represented as proportions in a mixture. There are some replicated mixtures so we create a distinct set of mixtures and average the outcome data across replicates.  

```{r chapter-10-data-prep}
data("concrete", package = "AppliedPredictiveModeling")

mixture_means <- 
 mixtures %>% 
 clean_names() %>% 
 group_by(cement, blast_furnace_slag, fly_ash, superplasticizer, 
          coarse_aggregate, fine_aggregate, water, age) %>% 
 summarize(
  compressive_strength = mean(compressive_strength), 
  .groups = "drop"
 )
```


```{r chapter-10-teardown, include = FALSE}
if (grepl("mingw32", R.Version()$platform)) {
 stopCluster(cl)
} 
```


