# Discriminant Analysis and Other Linear Classification Models

```{r chapter-12-startup, include = FALSE}
knitr::opts_chunk$set(fig.path = "figures/")
library(MASS)
library(mixOmics)
library(plsmod)
library(glmnet)
library(tidymodels)
library(workflowsets)
library(discrim)

cores <- parallel::detectCores()
if (!grepl("mingw32", R.Version()$platform)) {
 library(doMC)
 registerDoMC(cores = cores)
} else {
  library(doParallel)
  cl <- makePSOCKcluster(cores)
  registerDoParallel(cl)
}

source("extras/overlay_roc_curves.R")
```


The R packages used in this chapter are: `r pkg_text(c("tidymodels", "MASS", "discrim", "glmnet",
"mixOmics", "plsmod"))`. 

## Case Study: Predicting Successful Grant Applications

These data are supplied in the GitHub repository. The object `grants_testing` holds the test set while `grants_other` contains the data used for modeling and initial evaluation. 

We recreated the same split used in _APM_ here. However, the nature of this data splitting scheme is fairly irregular. For this reason, a non-standard `r pkg(rsample)` method was used; the results are the same as those produced by a more standard application of `initial_split()`. The code to create the `grants_split` object is found in [`extras/grants_splits.R`](https://github.com/topepo/tidy-apm/tree/main/extras/grants_splits.R).   

```{r chapter-12-data}
library(tidymodels)
tidymodels_prefer()
data(grants)

ls(pattern = "grants")

load("RData/grants_split.RData")

grants_split
nrow(grants_test)
```
`grants_split` is a validation-type resampling object. As noted in _APM_, $n = `r nrow(analysis(grants_split$splits[[1]]))`$ grants is use for model fitting while we predict $n = `r nrow(assessment(grants_split$splits[[1]]))`$ grants to measure model performance. We create the final model using both of these data sets. The final test set contains $n = `r nrow(grants_test)`$ grants are used as the final test set assessment.  


```{r chapter-12-teardown, include = FALSE}
if (grepl("mingw32", R.Version()$platform)) {
 stopCluster(cl)
} 
```


